---
title: "Sucholeiki_Robert_Code"
author: "Robbie Sucholeiki"
date: "2023-03-05"
output: html_document
---

# Investigating the Covid Response Retrospective Voting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load and Mutate Data
The following code reads in necessary packages and loads ANES data, selecting and cleaning relevant variables. It then mutates a recoded covid_approval score. We then mutate the following variables for use in our regresison analysis: a binary VoteBiden dependent variable and PartyID and gender covariates. 
```{r}
# Read packages
library(tidyverse)
library(stargazer)

# Load in and clean ANES data---------------------------------------------------
ANES <- read_dta("C:/Users/mysti/Desktop/SSI_II/ANES_2020/anes_timeseries_2020_stata_20220210.dta") %>%  
# Select variables of interest
  select(pid3 = V201228,
         pid_strong = V201229, 
         pid_lean = V201230,
         pid7 = V201231x,
         # covid approval
         covid_approval = V201144x,
         # candidate evaluations
         bidenft = V201151,
         trumpft = V201152,
         # vote choice
         votepres = V202110x,
         # Covariates
         age = V201507x,
         marital = V201508,
         educ = V201510,
         educ5 = V201511x,
         race = V201549x,
         gender = V201600,
         income = V201617x,
         ideology = V201200) %>%
  # Recode Missing Values as NA
  mutate(pid3 = ifelse(pid3 %in% c(-9, -8, -4, 0, 5), NA, pid3),
         pid_strong = ifelse(pid_strong < 0 , NA, pid_strong),
         pid_lean = ifelse(pid_lean < 0, NA, pid_lean),
         pid7 = ifelse(pid7 < 0, NA, pid7),
         covid_approval = ifelse(covid_approval < 0, NA, covid_approval),
         bidenft = ifelse(bidenft %in% c(-9, -4, 998), NA, bidenft),
         trumpft = ifelse(trumpft %in% c(-9, -4, 998), NA, trumpft),
         age = ifelse(age < 0, NA, age),
         marital = ifelse(marital < 0, NA, marital),
         educ = ifelse(educ %in% c(-9, -8, 95), NA, educ),
         educ5 = ifelse(educ5 < 0, NA, educ5),
         race = ifelse(race < 0, NA, race),
         gender = ifelse(gender < 0, NA, gender),
         income = ifelse(income < 0, NA, income),
         ideology = ifelse(ideology %in% c(-9, -8, 99), NA, ideology))


# Mutate a covid_retro score for Trump's covid response-------------------------
# covid_retro ranges from -1 ("disapprove strongly") to 2 ("approve strongly")
ANES2 <- ANES %>%
  # filter party ID and covid_approval
  filter(pid3 %in% c(1:3),
         is.na(covid_approval) == FALSE) %>%
  # mutate covid_retro score
  mutate(covid_retro = case_when(covid_approval == 1 ~ 1,
                                 covid_approval == 2 ~ 2/3,
                                 covid_approval == 3 ~ 1/3,
                                 covid_approval == 4 ~ 0))
         

#Examine covid_retro in a table-------------------------------------------------
table(ANES2$covid_retro)


# Filter and mutate additional variables for use in regression
ANES3 <- ANES2 %>%
  #filter relavent variables
  filter(votepres %in% c(1:2), #filter only Biden and Trump votes
         #filter relavent covariates
         pid7 %in% c(1:7), 
         gender %in% c(1:2), 
         is.na(age) == FALSE, 
         is.na(income) == FALSE, 
         is.na(educ5) == FALSE,
         is.na(income) == FALSE) %>% 
  #mutate voteBiden, partyID, and create a dummy variable for gender
  mutate(voteBiden = case_when(votepres == 2 ~ 0, #assign 0 for Trump votes
                               votepres == 1 ~ 1),#assign 1 for Biden votes
         partyID = (pid7 - 4) / 3, #mutate party ID from -1 (strong Dem) to 1 (strong Rep)
         female = ifelse(gender == 2, 1, 0), #dummy variable for gender, 1 = female)
         education = (educ5 - 3)/2, #mutate education from -1 (< high school) to 1 (graduate school)
         Democrat = ifelse(pid3 == 1, 1, 0), #dummy variable for Democrat Party ID
         Republican = ifelse(pid3 == 2, 1, 0),#dummy variable for Republican Party ID
         Independent = ifelse(pid3 == 3, 1, 0)) #dummy variable for Independent party ID
  

```


## Visualize Data
The following code filters Biden thermometer scores, then plots Biden thermometer score against covid_retro, stratifying by Party ID.
```{r}
scatterplot_stratified <- ANES2 %>%
   filter(bidenft %in% c(0:100)) %>%
   mutate(Party = case_when(pid3 == 1 ~ "Democratic",
                            pid3 == 3 ~ "Independent",
                            pid3 == 2 ~ "Republican",)) %>%
   ggplot(aes(x = covid_retro, y = bidenft)) + #plot bidenft vs covid_retro
   geom_point(size = 2, alpha = .5, shape = 1) + #scatterplot with adjusted parameters
   geom_smooth(method = "lm", size = 2, color = "red", se = F) + #adjust trendline
   scale_x_continuous(breaks = c(0, 1/3, 2/3, 1), 
                      labels = c("0", "1/3", "2/3", "1")) + #custom x axis breaks
   facet_wrap(~ Party, ncol = 3) + #facet_rap to show patterns by party id
   labs(x = "Trump COVID-19 Response Approval (Low <-> High) \n 
            0 = Strong Disapproval | 1/3 = Weak Disapproval | 2/3 = Weak Approval | 1 = Strong Approval", #apply custom axes labels
       y = "Feeling Thermometer Toward Biden", 
       title = "Feeling Thermometer Toward Biden vs Trump COVID-19 Response Approval") +
  guides(x =  guide_axis(angle = 0)) +
  theme_bw()

scatterplot_stratified
```
The following code filters Biden thermometer scores, then plots Biden thermometer score against covid_retro and applies a regression line by party ID
````{r}
scatterplot <- ANES3 %>%
  filter(bidenft %in% c(0:100)) %>%
  mutate(Party = case_when(pid3 == 1 ~ "Democratic",
                           pid3 == 3 ~ "Independent",
                           pid3 == 2 ~ "Republican")) %>%
  ggplot(aes(x = covid_retro, y = bidenft, color = Party)) + #plot bidenft vs covid approval
  geom_jitter(alpha = .75, shape = 1) + #jitter scatterplot with adjusted parameters
  geom_smooth(method = "lm", size = 2, se = F) + #adjust trendline
  scale_x_continuous(breaks = c(0, 1/3, 2/3, 1), #custom x axis breaks
      labels = c("Disapprove Strongly", 
                 "Disapprove Weakly", 
                 "Approve Weakly",
                 "Approve Strongly")) +
  labs(x = "Trump COVID-19 Response Approval (Low <-> High)", #apply custom axes labels
       y = "Feeling Thermometer Toward Biden", 
       title = "Feeling Thermometer Toward Biden vs Trump COVID-19 Response Approval") +
  theme_bw() + 
  scale_color_manual(values = c("#276DC2", "grey", "firebrick2"))

scatterplot

```
The following code produces a histogram of Trump Covid Response Approval Ratings
```{r}
histogram_covid <- ANES2 %>%
  mutate(Party = case_when(pid3 == 1 ~ "Democratic",
                           pid3 == 3 ~ "Independent",
                           pid3 == 2 ~ "Republican")) %>%
  mutate(covid_scale = 4 - covid_approval) %>%
  ggplot(aes(x = covid_scale, fill = Party)) + #histogram of variable covid_retro
  geom_histogram(binwidth = 0.5) + #adjust color and binwidth
  scale_x_continuous(breaks = c(0, 1, 2, 3), #apply custom x axis breaks
                     labels = c("Disapprove Strongly", 
                                "Disapprove Weakly", 
                                "Approve Weakly", 
                                "Approve Strongly")) +
  labs(x = "Trump COVID-19 Response Approval (Low <-> High)", #apply custom axes labels
       y = "Number of Respondents", 
       title = "Histogram of Trump COVID-19 Response Approval in ANES 2020") +
  theme_bw() + 
  scale_fill_manual(values = c("#276DC2", "grey", "firebrick2"))

histogram_covid
```
The following code produces a histogram of vote choice for Biden and Trump
```{r}
histogram_vote <- ANES2 %>%
  filter(votepres %in% c(1:2)) %>% #filter only Biden and Trump votes)
  mutate(Party = case_when(pid3 == 1 ~ "Democratic",
                           pid3 == 3 ~ "Independent",
                           pid3 == 2 ~ "Republican")) %>%
  mutate(voteBiden = case_when(votepres == 2 ~ 0, #assign 0 for Trump votes
                               votepres == 1 ~ 1)) %>% #assign 1 for Biden votes
  ggplot(aes(x = voteBiden, fill = Party)) + #histogram of variable covid_retro
  geom_histogram(binwidth = 0.5) + #adjust color and binwidth
  scale_x_continuous(breaks = c(0,1), #apply custom x axis breaks
                     labels = c("Voted for Trump", 
                                "Voted for Biden")) +
  labs(x = "Reported Vote", #apply custom axes labels
       y = "Number of Respondents", 
       title = "Histogram of Votes for Biden and Trump in ANES 2020") +
  theme_bw() + 
  scale_fill_manual(values = c("#276DC2", "grey", "firebrick2"))

histogram_vote
```



## Linear Probability Models

Run a linear probability model with covid_retro as the explanatory variable and voteBiden as the dependent variable
```{r}
lpm1 <- lm(voteBiden ~ covid_retro, data = ANES3)
summary(lpm1)  
```

Run a linear probability model with party ID as a control variable
```{r}
lpm2 <- lm(voteBiden ~ covid_retro + partyID, data = ANES3)
summary(lpm2)  
```
Run a linear probablility model with all control variables
```{r}
lpm3 <- lm(voteBiden ~ covid_retro + partyID + age + gender + education + income, data = ANES3)
summary(lpm3)  
```
Run a linear probablility model with interaction terms for party ID
```{r}
lpm4 <- lm(voteBiden ~ covid_retro * Democrat + covid_retro * Republican, data = ANES3)
summary(lpm4)  
```

Run a linear probability model with interaction terms for party ID and control variables
```{r}
lpm5 <- lm(voteBiden ~ covid_retro * Democrat + covid_retro * Republican + age + gender + education + income, data = ANES3)
summary(lpm5)  
```


```{r}
stargazer(lpm1, lpm2, lpm3,
          type = "text", 
          omit.stat = c("adj.rsq", "f", "ser"),
          title = "Figure 1: Trump COVID-19 Response Approval And Vote Choice in the 2020 Election",
          dep.var.labels = c("Vote Choice Toward Biden"),
          covariate.labels = c("Trump COVID-19 Response Approval",
                                "Party ID (-1 Dem; 0 Ind; 1 Rep)",
                                "Age", "Gender", "Education Level", "Income"),
          out = "C:/Users/mysti/Desktop/SSI_II/Final_Assignment/lm_reg_table.html")
```

```{r}
stargazer(lpm4, lpm5,
          type = "text", 
          omit.stat = c("adj.rsq", "f", "ser"),
          title = "Figure 2: Vote Choice in the 2020 Election: Trump COVID-19 Response Approval 
          and Interactions with PartyID",
          dep.var.labels = c("Vote Choice Toward Biden"),
          covariate.labels = c("Trump COVID-19 Response Approval",
                                "Democrat", "Republican",
                                "Age", "Gender", "Education Level", "Income", 
                                 "Democrat COVID-19 Interaction",
                                 "Republican COVID-19 Interaction"),
          out = "C:/Users/mysti/Desktop/SSI_II/Final_Assignment/lm_interaction_reg_table.html")

```


